---
title: "STAA 566 Dynamic Figure"
author: "Delsie Johnson"
format: html
html:
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE)
```

# 2022 WNBA Playoffs: Dynamic Figure

**Data source:** "WNBA_playoffs.csv" from https://stats.wnba.com/players/boxscores-traditional/. csv file included

**What I want to convey:** I want to convey the points in each playoff game for each player of the two WNBA teams who made it to the championship. I want to show how each player varied in points per game and also how each team varied compared to the other team. I want this graph to give insight into how players points in a game changed during the playoffs and if it had any influence on which team won overall. Note: the Las Vegas Aces won the 2022 WNBA championship over the Connecticut Sun.

**Functionality/formatting:** I allow the user to hover over the points in the graph. I do this so they can see more information including the team, player, points, and game date. The user can also hide/show each team. This allows them to only look at the scoring data for one team at a time, which makes the graph less crowded and easier to see if that is what the user wants.

## R Code:

```{r}
# load libraries
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(plotly)
library(htmlwidgets)

# read in data from stats.wnba.com
WNBA <- read.csv("WNBA_playoffs.csv", header=T)
WNBA$game_date <- as.Date(WNBA$game_date, format="%m/%d/%y")
# renaming variables for formatting
WNBA$Points <- WNBA$pts
WNBA$Game.Date <- WNBA$game_date
WNBA$Player <- WNBA$player
WNBA$Team <- WNBA$team

# create ggplot
p_playoffs <- ggplot(data=WNBA, aes(x=Game.Date, y=Points, col=Team)) +
  ### plotting Conneticut data
  geom_point(data=WNBA %>% filter(Team=="CON"), 
             aes(group=Player), 
             position = position_jitter(w = 0.2, h = 0), 
             alpha=0.6)+
  # connecting lines of top 2 scorers for Conneticut
  geom_line(data=WNBA %>% filter(Player=="JONQUEL JONES"),
            linetype="dashed")+
  geom_line(data=WNBA %>% filter(Player=="ALYSSA THOMAS"),
            linetype="dotted")+
  
  ### plotting Las Vegas data
  geom_point(data=WNBA %>% filter(Team=="LVA"),
             aes(group=Player), 
             position = position_jitter(w = 0.2, h = 0),
             alpha=0.6)+
  # connecting lines of top 2 scorers for Las Vegas
  geom_line(data=WNBA %>% filter(Player=="CHELSEA GRAY"),
            linetype="dashed")+
  geom_line(data=WNBA %>% filter(Player=="A'JA WILSON"), 
            linetype="dotted")+
  labs(x="",
       y="Points", 
       title="2022 WNBA Playoffs: Las Vegas & Conneticut Points by Players")+
  theme_classic()+
  theme(
        panel.grid.major.y = element_line(linetype = 3, color="grey80"))

# make ggplot dynamic with plotly
p_playoffs_plotly <- ggplotly(p_playoffs)
p_playoffs_plotly <- p_playoffs_plotly %>% # adding subtitle using HTML to format text
  layout(title = list(text = paste0("2022 WNBA Playoffs: Las Vegas & Conneticut Points by Players",
                                    '<br>',
                                    '<sup>',
                                    "Colored lines represent scoring trends by Jonquel Jones, Alyssa Thomas, Chelsea Gray & A'ja Wilson",
                                    '</sup>')))
```

## Save Figure as PDF

```{r save as pdf}
pdf("WNBA_playoffs.pdf", heigh=6, width=8)
print(p_playoffs_plotly)
dev.off()
```

## Display Dynamic Figure in HTML:

```{r display}
p_playoffs_plotly
```

**Note:** Jonquel Jones and Alyssa Thomas were Connecticut's top two scorers by points per game over the course of the playoffs. Chelsea Gray and A'ja Wilson were Las Vegas' top two scorers by points per game over the course of the playoffs.
